---
title: "White Wine Analysis"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

---
title: "White Wine Quality Analysis"
editor_options:
  chunk_output_type: console
output:
  html_document: 
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    df_print: paged
  pdf_document: default
---



# Disclaimer :  
### This tidy data set contains 4,898 white wines with 11 variables on quantifying the chemical properties of each wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent). In This project we want to discover what is chemical properties affect white wine quality.


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.
library(ggplot2)
library(grid)
library(reshape2)
```

```{r echo=FALSE, Load_the_Data}


# Load the Data: 


setwd('C:/Users/farah/OneDrive/Desktop')

wqw <- read.csv('wqw.csv')


```


# Univariate Plots Section


```{r echo=FALSE, Load_Data}

# Load the Data
nrow(wqw)
ncol(wqw)
head(wqw, n=10)
tail(wqw, n=8)
summary(wqw)
str(wqw)
```



```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}


# explore fixed acidity ,volatile acidity, and citric acid of wine
summary(wqw$fixed.acidity)
summary(wqw$volatile.acidity)
summary(wqw$citric.acid)


```

**As we can see from the summary fuction, Violate acidity, and citric acid recorded the lowest amont compared to fixed acidity **
**The fixed acidity recorded mean with 6.855 approximatily, and maximum level with 14, which data ranged between 3 to 14.**
**The volatile acidity recorded mean with 0.278 approximatily, and maximum level with 1.1000, which data ranged between 0.0800 to 1.1000.**
**The citric acid recorded mean with 0.334 approximatily, and maximum level with 1.6600, which data ranged between 0.0000 to 1.6600.**

**We still need to discover more to figure out  the quality of white wine.**


```{r echo=FALSE, message=FALSE, warning=FALSE}


#plot the fixed acidity chart
ggplot(aes(x = fixed.acidity), data = wqw) +
  geom_histogram(binwidth = 0.25,color = 'black', fill = 'deepskyblue4') +
  scale_x_continuous(breaks = seq(4, 12, 1), limits = c(4, 12)) + 
  xlab("Fixed Acdity")+
  ylab("frequency")+
  ggtitle("White Wine Fixed Acdity")

#plot the volatile acidity chart

ggplot(aes(x = volatile.acidity), data = wqw) +
  geom_histogram(binwidth = 0.05,color = 'black', fill = 'cadetblue2') +
  scale_x_log10(breaks = seq(0.05, 0.5, 0.05), limits = c(0.05, 0.5)) + 
  xlab("volatile Acdity")+
  ylab("frequency")+
  ggtitle("White Wine volatile Acdity")

#plot the citric acid chart

ggplot(aes(x = citric.acid), data = wqw) +
  geom_histogram(binwidth = 0.05,color = 'black', fill = 'lightpink') +
  scale_x_log10(breaks = seq(0.05, 0.5, 0.05), limits = c(0.05, 0.5)) +
  xlab("citric Acid")+
  ylab("frequency")+
  ggtitle("White Wine citric acid")

#grid.arrange(p1,p2,p3, ncol=1)

```

**From the previous charts , Fixed acidity, volatile acidity and citric acidity have a normally distributed and left skewed shaped, "except the shape on 0.05's spike on Citric acid"**


```{r echo=FALSE, message=FALSE, warning=FALSE}


#plot the ph of white wine 
ggplot(aes(x = pH), data = wqw) +
  geom_histogram(binwidth = 0.05,color = 'black', fill = 'gray58') +
  scale_x_log10()+  xlab("pH of wine")+
  ylab("frequency")+
  ggtitle("White Wine pH")

#plot the sulpates of white wine
ggplot(aes(x = sulphates), data = wqw) +
  geom_histogram(binwidth = 0.05,color = 'black', fill = 'firebrick4') +
  scale_x_log10() +  xlab("Sulphates of wine")+
  ylab("frequency")+
  ggtitle("White Wine Sulphates")


#plot residual of sugar
ggplot(aes(x = residual.sugar), data = wqw) +
  geom_histogram(binwidth = 0.05,color = 'black', fill = 'gold2') +
  scale_x_log10() +  xlab("residual sugar")+
  ylab("frequency")+
  ggtitle("White Wine residual sugar")

#(p1_1,p1_2,p1_3, ncol=1)


```

**From the previous charts , pH rate, sulphates rate and residual sugar we can notes that pH of white wine have highest number with 4000 on 3 of pH, and sulphates rate we notes that white wine recorded highest rate with 1000 on 0.5 of sulphates.Finally, residual of sugar we can notes sudenly increasing and dicreading.**



# Univariate Analysis



### What is the structure of your dataset?
**Our dataset encompass 4899 observations and 13 variables,We have 11 chemical properties :fixed acidity, volatile acidity, Citric Acid , Residual Sugar   , Chloride ,Free Sulfur Dioxide, Sulfur dioxide, Density, pH, Dulphates, Alcohol (g / dm^3).**

### What is/are the main feature(s) of interest in your dataset?
**I am interested in discovering the quality and the rate of chloride and pH of white wine.**

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
**Type of acidity such citric and volatile acid**


### Did you create any new variables from existing variables in the dataset?
**Yes,  I found that the rate with numeric values is not make a sense, so I created new variable with: very bad from 1 to 3 and good for 4 and 5, 6 and 7 is recorded as very good, and from 8 to 10 recorded as the excellent quality of white wine**

```{r echo=FALSE, Univariate_Plots}
#create new column(quality_rate)
wqw$Quality_rate <- factor(wqw$quality,
levels = c(1,2,3,4,5,6,7,8,9,10),
labels = c("ver bad", "very bad", "very bad", "good", "good", "very good",
           "very good", "excellent", 'excellent', 'excellent')) 

table(wqw$quality)

wqw$quality_rate<-cut(wqw$quality, c(1,3,4,5,6,7,8,10))

```

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
**I did not find unusual content on the charts.**

# Bivariate Plots Section







```{r echo=FALSE, message=FALSE, warning=FALSE}


summary(wqw$free.sulfur.dioxide)

#plot the rate of free sulfur dioxide
ggplot(aes(x=free.sulfur.dioxide, y=total.sulfur.dioxide), data =wqw)+
  geom_point(color="#20b2aa")+ xlim(0, quantile(wqw$free.sulfur.dioxide,0.95))+
  ylim(0 ,quantile(wqw$total.sulfur.dioxide, 0.95))+
  geom_smooth(method = 'lm', color='red')+
  ggtitle("the rate of free sulfur dioxide on total sulfur dioxide for whie wine")
  
cor.test(wqw$free.sulfur.dioxide, wqw$total.sulfur.dioxide)

```

**In this scatter chart we want to discover the relationship among free sulfur dioxide and total sulfur dioxide, as we can see after calculating the correlation=0.615501, which is higher than 0.5, so it is modarate relation.**

# Bivariate Analysis


### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

**From the previous charts we discovered many properties of white wine:**

- **The quality rate of alcohol and pH rate, , from the chart we can notes that very bad white wine recorded the highest rate of pH on 9 to 10 of alcohol rate, then good rate is became second with 3.5 of pH in 11 rate of alcohol.**

- **For the second chartwe found the grand mean, the very good white wine is closer to mean.**

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?
- **The relationship of sulfur content have the moderate relatins.**

### What was the strongest relationship you found?

**I will discover it for the next part**

# Multivariate Plots Section



```{r echo=FALSE, message=FALSE, warning=FALSE}



qplot(data=wqw, x=alcohol, y=sulphates, color=quality_rate,
      size=I(5), alpha=I(0.15))+
  geom_smooth(method = 'lm', color='red')+ 
  ggtitle("alcohol and sulphates rate for white wine types") +
  xlab("alcohol g / dm^3")+ylab("sulphates g / dm^3")
                                                            

cor.test(wqw$alcohol, wqw$sulphates)



```

**we can illustrate that, the relationship among alcohol and sulphates for each type of white wine have a negative weak relation with -0.0174 correlation.**




```{r echo=FALSE, message=FALSE, warning=FALSE}

p6<-ggplot(data=wqw, aes(pH, volatile.acidity,color=quality_rate))
p6+geom_boxplot(size=2.5, alpha=1/20)+
  xlab("pH g / dm^3 ")+
  ylab("volatile acidity g / dm^3 ")+
  ggtitle("volatile acidity and pH rate on each type of white wine")

```

**In this part we want to find the highest and lowest rate of volatile acidity form white wine type,as we can see from the chart that bad type (3-4) of white wine recorded the highest rate of volatile acidity, then very bad type (1-3) , finally very good type of white wine(5-6), recorded the lowest rate.**

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqw[wqw$alcohol<13,], aes(x=alcohol, y=pH, color=quality_rate))+
  geom_point(alpha=0.5) +
  geom_smooth(method = 'lm', color='red')+ 
   xlab("alcohol g / dm^3 ")+
  ylab("pH g / dm^3 ")+
  ggtitle("the ralatonship among alcohol and pH rate of white wine")
cor.test(wqw$alcohol, wqw$pH)
```

**in this chart we want to descover the relationship among pH and alcohol rate on ewach type of white wine, and after calculating the correlation we can see that 0.1214 is less that 0.3 which is weak ralationship**

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqw[wqw$citric.acid<1.5,], aes(x=citric.acid,y=pH,color=quality_rate))+
  geom_point(alpha=0.5) +
  geom_smooth(method = 'lm', color='red')+
  geom_point(alpha=0.75)+geom_smooth(method = 'lm', color='red')+
   xlab("citric acid g / dm^3 ")+
  ylab("pH g / dm^3 ")+
  ggtitle("the ralatonship among citric acid and pH rate of white wine")
cor.test(wqw$citric.acid, wqw$pH)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqw, aes(x=volatile.acidity, y=pH, color=quality_rate))+
  geom_point(alpha=0.75)+ 
  geom_smooth(method = 'lm', color='red')+geom_point(alpha=0.75)+
  geom_smooth(method = 'lm', color='red')+
  xlab("volatile acidity g / dm^3 ")+
  ylab("pH g / dm^3 ")+
  ggtitle("the ralatonship among volatile acidity and pH rate of white wine")
cor.test(wqw$volatile.acidity, wqw$pH)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=wqw[wqw$chlorides<0.20,], aes(x=chlorides, y=pH, color=quality_rate))+
  geom_point(alpha=0.75)+ geom_smooth(method = 'lm', color='red')+
  xlab("chlorides g / dm^3 ")+
  ylab("pH g / dm^3 ")+
  ggtitle("the ralatonship among cholorides acidity and pH rate of white wine")

cor.test(wqw$chlorides, wqw$pH)
```

**As we can see from the previous charts there are negative relations among citric.acid, fixed.acidity, volatile.acidity, chlorides and pH, which is less than 0.3, in which it is considered as a weak relation.**

#Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

**As we can see from the previous charts there are negative relations among citric.acid, fixed.acidity, volatile.acidity, chlorides and pH, which is less than 0.3, in which it is considered as a weak relation.**

### Were there any interesting or surprising interactions between features?

**Before starting analysis I guessed that the relations among acidity and pH will be string relations, after calculating the correlations and displaying the charts I find it as a weak relations.**


------

# Final Plots and Summary

### Plot One


```{r echo=FALSE, plot_one}



ggplot(data=wqw[wqw$chlorides<0.15,], aes(x=chlorides, y= residual.sugar,color=quality_rate))+
  geom_point(alpha=0.75)+ geom_smooth(method = 'lm', color='red')+
  scale_y_log10()+
  ylab("residual sugar g / dm^3")+
  xlab("cholorides g / dm^3")+
  ggtitle("the ralatonship among cholorides acidity and pH rate of white wine")

cor.test(wqw$chlorides, wqw$residual.sugar)


```


### Description One
**The correlation ais equal 0.088 which is less than 0.3, the the relations among chlorides and residual of sugar is very weak relations**

### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(data=wqw[wqw$alcohol<13,], aes(x=alcohol, y=pH, color=quality_rate))+
  geom_point(alpha=0.5) +
  geom_smooth(method = 'lm', color='red')+
  xlab("alcohol g / dm^3")+
  ylab("pH g / dm^3")+
  ggtitle("the ralatonship among alcohol and pH rate of white wine")
cor.test(wqw$alcohol, wqw$pH)

```

### Description Two
The correlation here is equal 0.1214, which less than 0.3 and considered as the weak relation among alcohol and pH rates.

### Plot Three
```{r echo=FALSE, Plot_Three}
p6<-ggplot(data=wqw, aes(pH, volatile.acidity,color=quality_rate))
p6+geom_boxplot(size=2.5, alpha=1/20)+
  xlab("pH g / dm^3")+
  ylab("volatile acidity g / dm^3")
  ggtitle("volatile acidity and pH rate on each type of white wine")

```

### Description Three

in this part we found that bad type(3-4) of white wine recorded the highest rate of volatile acidity, then very bad type (1-3) , finally very good type of white wine(5-6), recorded the lowest rate.

## Reflection

### Did you meet any challenge during your analysis?
yes because when I use this data I want to find the effect and harms of drinking wine.

###what information do you need to improve your analysis?
it wil be better if the data included the effect of wine then the analysis will improving and fining out many results.

------
